---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# handbaloner
- Documentación en [español](README.es.md)
<!-- badges: start -->
<!-- badges: end -->

The `handbaloner` packages has useful function for the visualization of handball
data. 

## Package installation

The development version can be installed from [GitHub](https://github.com/)
with:

``` r
# install.packages("devtools")
devtools::install_github("telaroz/handbaloner")
```
## Court visualization examples

In this example, we can draw a basic court.

```{r basic example}
library(handbaloner)

court()
```

We can change the colours, rotate vertically and mirror the court with the 
functions' parameters

```{r example with colours}
court(vertical = TRUE, flip = TRUE, court_color = "orange", 
      area_color = "#3431A2", lines_color = "black")
```

As the plots are generated with ggplot, we can describe the colourse with their
HEX code, rgb, number or by its name in english; see: [use of colours in ggplot2](https://r-graph-gallery.com/ggplot2-color.html)

We can also draw half a court

```{r half court example}
half_court(vertical = TRUE, court_color = colors()[36], 
      area_color = rgb(red = 0.2, green = 0.4, blue = 0.6), 
      lines_color = "yellow")
```

Another useful function is `distance_to_goal` which measures the distance from a
point of the field to its closest goal, given some coordinates ([-40, 40] in the
x axis and [-20, 20] in the y axis): 

```{r}
distance_to_goal(c(10, 3))
```

Also, as this is a ggplot, we can add some additional layers. For example, let's
generate a `data.frame` with some coordinates of shots and whether they were
goals or not. We will add the distance to goal as a column and plot in green and
red if the shots were goal or not. 


```{r}
tiros <- dplyr::tibble(x = c(-13, -12, 11, -11, 9.5),
                       y = c(2, 5, -3, -1, 0),
                       gol = c(1, 0, 1, 1, 0))

dplyr::mutate(tiros, distance_to_goal = purrr::map2_dbl(x, y, ~ distance_to_goal(c(.x, .y))))

```

```{r court with shots}
court() +
  ggplot2::geom_point(data = tiros, ggplot2::aes(x, y),
                      color = ifelse(tiros$gol == 1, 'Green', 'Red'),
                      size = 4)
```

## Generate Play by Play tidy Data from IHF files

First, you need  to download the PBP pdf file. You can use the `scrape_from_ihf`
function to do so. Find the link for the match information and set the folder to
download the file. 

For the first match of the 2023 World Men's Handball Championship, you can 
download all PDFs as follows:

```{r download files, eval = FALSE}

scrape_from_ihf(link = "https://www.ihf.info/competitions/men/308/28th-ihf-men039s-world-championship-2023/101253/match-center/118895",
                folder = "pol_swe")
```

Now, use the `generate_tidy_pbp` to generate a `data.frame` in a tidy format. 

```{r generate tidy pbp}
generate_tidy_pbp("pol_swe/02PBP.PDF")
```

