---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# handbaloner
- Documentación en [español](README.es.md)
<!-- badges: start -->
<!-- badges: end -->

The `handbaloner` packages has useful function for the visualization of handball
data. 

## Package installation

The development version can be installed from [GitHub](https://github.com/)
with:

``` r
# install.packages("devtools")
devtools::install_github("telaroz/handbaloner")
```
## Court visualization examples

In this example, we can draw a basic court.


```{r basic example}
library(handbaloner)

court()
```

Podemos cambiar los colores, rotar verticalmente y en espejo con sus parámetros.

We can change the colours, rotate vertically and mirror the court with the 
functions' parameters

```{r example with colours}
court(vertical = TRUE, flip = TRUE, court_color = "orange", 
      area_color = "#3431A2", lines_color = "black")
```

As the plots are generated with ggplot, we can describe the colourse with their
HEX code, rgb, number or by its name in english; see: [use of colours in ggplot2](https://r-graph-gallery.com/ggplot2-color.html)

We can also draw half a court

```{r half court example}
half_court(vertical = TRUE, court_color = colors()[36], 
      area_color = rgb(red = 0.2, green = 0.4, blue = 0.6), 
      lines_color = "yellow")
```

Another useful function is `distance_to_goal` which measures the distance from a
point of the field to its closest goal, given some coordinates ([-40, 40] in the
x axis and [-20, 20] in the y axis): 

```{r}
distance_to_goal(c(10, 3))
```

Also, as this is a ggplot, we can add some additional layers. For example, let's
generate a `data.frame` with some coordinates of shots and whether they were
goals or not. We will add the distance to goal as a column and plot in green and
red if the shots were goal or not. 


```{r}
tiros <- dplyr::tibble(x = c(-13, -12, 11, -11, 9.5),
                       y = c(2, 5, -3, -1, 0),
                       gol = c(1, 0, 1, 1, 0))

dplyr::mutate(tiros, distance_to_goal = purrr::map2_dbl(x, y, ~ distance_to_goal(c(.x, .y))))

```

```{r court with shots}
court() +
  ggplot2::geom_point(data = tiros, ggplot2::aes(x, y),
                      color = ifelse(tiros$gol == 1, 'Green', 'Red'),
                      size = 4)
```

